Spatial Reasoning and Statistical Graphics

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Citations for Mahbub and Niladri's arxiv papers.

Statistical lineups~\cite{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition). Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation). By breaking the lineup task down into component parts, we may be able to determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability. These assessments are then used to account for some of the variation in lineup performance differences. In addition, we also examine the effect that previous experiences (science-based major, research experience, Auto-CAD skills) have on lineup performance and the implications for the use of lineups to assess the graphical display of statistical information.

>>>>

Statistical lineups~\cite{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition). Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation). By breaking the lineup task down into component parts, we may be able to determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability. \newtext{Demographic factors are known to contribute to lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender affected the amount of time spent on lineups \cite{humanfactorslineups}.}\newtext{In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.}\newtext{In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability, with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance. }
